Supplementary Information

### 2020-11-07 doc/manuscript/MEE_manuscript_SI.tex
replace ∆r with RW, add rows to table
in the meantime, doc stopped knitting (not because of these changes)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplementary Information},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

@@

In either case we need bark thickness--ideally allometries describing
the relationship between DBH and bark thickness (Table S4). This is
especially critical for thick-barked species When bark thickness data
were available, we generated allometries \ldots{} lognormal model with
intercept forced to zero: lm(bark\_depth.mm \textasciitilde{} -1 +
log(dbh\_no\_bark.cm+1):bark\_species, data = bark)
(\href{https://github.com/EcoClimLab/ForestGEO_dendro/issues/8}{issue
\#8 in ForestGEO\_dendro})

@@

(**\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/35}{ISSUE
\#35 in ForestGEO-climate-sensitivity}

@@

Our analysis included two sites where climate change has had pronounced
effects on tree growth: Scotty Creek, NW Territories, Canada (SC) and
Little Tesuque, New Mexico, USA (LT). At SC, {[}temperatures have
increased by X \(^\circ\) over X years{]}\ldots, resulting in negative
growth trends in basal area index (\(BAI\)) starting around 1950 and
significant growth declines since 1970 in 56\% of trees
{[}@sniderhan\_growth\_2016{]}.

@@

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{l>{\raggedright\arraybackslash}p{3cm}rr>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{1.5cm}llll}
\toprule
site code & site name & latitude & longitude & cores within ForestGEO plot? & canopy positions & tree statuses & date range & dormant season & months in climwin\\
\midrule
BCI & Barro Colorado Island & 9.15430 & -79.8461 & no & canopy & live, dead & 1931-2014 & Nov-Apr & pOct-cDec\\
HKK & Huai Kha Khaeng & 15.63240 & 99.2170 & no & all & live & 1903-2011 &  & pOct-cDec\\
LT & Little Tesuque & 35.73838 & -105.8382 & n.a. & all & live & 1903-2018 &  & pMay-cAug\\
CB & Utah Forest Dynamics Plot & 37.66150 & -112.8525 &  &  &  & 1903-2007 &  & pMay-cAug\\
SCBI & Smithsonian Conservation Biology Institute & 38.89350 & -78.1454 & yes & all & live, dead & 1903-2017 & Oct-Apr & pMay-cAug\\
\addlinespace
LDW & Lilly Dickey Woods & 39.23590 & -86.2181 &  & canopy & live, dead & 1903-2019 &  & pMay-cAug\\
HF & Harvard Forest & 42.53880 & -72.1755 & yes & all & live, dead & 1903-2014 &  & pMay-cAug\\
NB & Niobrara/Hansley & 42.78000 & -100.0210 &  &  &  &  &  & pMay-cAug\\
ZOF & Zofin Forest Dynamics Plot & 48.66380 & 14.7073 & some & all &  & 1903-2013 &  & pMay-cAug\\
SC & Scotty Creek & 61.30000 & -121.3000 & no & all & live, dead & 1903-2013 &  & pMay-cAug\\
\bottomrule
\end{tabular}}
\end{table}

@@

(**\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/72}{ISSUE
\#72 in ForestGEO-climate-sensitivity}

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{llllllll}
\toprule
species code & family & latin name & site(s) sampled & leaf type & leaf phenology & light requirements & bark allometry*\\
\midrule
$species_code in sitespecies.csv$ & family in sitespecies.csv & $latin_name$ in sitespecies.csv & site codes & $leaf_type$ in sitespecies.csv & $leaf_phenology$ in sitespecies.csv & $light_requirements$ in sitespecies.csv & "$bark_species$ in $bark_site$"\\
\bottomrule
\end{tabular}}
\end{table}

@@

(**\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/73}{ISSUE
\#73 in ForestGEO-climate-sensitivity}

@@

(**\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/71}{ISSUE
\#71 in ForestGEO-climate-sensitivity}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplementary Information},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

@@

\emph{This is still rough/ mostly notes.}

@@

In either case we need bark thickness--ideally allometries describing
the relationship between DBH and bark thickness (Table S4). This is
especially critical for thick-barked species. When bark thickness data
were available, we generated allometries
(\href{https://github.com/EcoClimLab/ForestGEO_dendro/issues/8}{issue
\#8 in ForestGEO\_dendro})\ldots{} lognormal model with intercept forced
to zero:
\texttt{lm(bark\_depth.mm\ \textasciitilde{}\ -1\ +\ log(dbh\_no\_bark.cm+1):bark\_species,\ data\ =\ bark)}.
When bark thickness data were not available, we used published bark
allometries from other sources (Table S4)

@@

\emph{(\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/35}{ISSUE
\#35 in ForestGEO-climate-sensitivity})}

To verify that our methods gave similar results to traditional methods,
we conducted qualitative comparisons of our results to previous studies
based on the same cores (Table \textbf{S5}). We also conducted a formal
comparison using identical tree-ring and climate data for four
well-studied species: PSME (Cedar Breaks, Utah), ABAL (Zofin), PIMA
(Scotty Creek), and LITU (SCBI) (Fig. \textbf{S1}). We compared results
from an analysis using conventional methods, as detailed below, to an
analysis using our method as described in the Methods section, but with
the \emph{climwin} climate variable selection process limited to just
the species of interest (as opposed to all species at the site), climate
variables considered individually rather than additively, and with start
date adjusted to match the conventional method.

The ring-width series from each core was standardized via ARSTAN using a
2/3rds \(n\) spline, where \(n\) is the number of years in the series
(\emph{Cook, 1985; Cook \& Kairiukstis, 1990- citations in Helcoski}).
\emph{(The following italic text is plagarized from Helcoski and needs
to be reworded:)} \emph{The influence of outliers in all series was
reduced using the adaptive power transformation, which also stabilises
the variance over time (Cook \& Peters, 1997). Next, each series was
stabilised using either the average correlation between raw ring-width
series (rbar) method or a 1/ 3rds spline method to adjust changes in
variance as series replication decreased towards the earlier portion of
each chronology (Jones et al., 1997). The 1/3rds spline method was
chosen when replication in the inner portion of each chronology (c.~the
inner 30--50 yr of each record depending on full chronology length)
dropped below three trees. Once that step was complete, a robust
biweight mean chronology for each species was calculated from the
ring-width indices (Cook, 1985). We chose to use residual chronologies
because the autoregressive standardisation process in creating them
removes much of the tree-level autocorrelation in growth and these
chronologies would most likely contain the most conservative information
on drivers of interannual growth (Cook, 1985).}

Following Helcoski et al. (2019), we defined chronology start dates
according to the subsample signal strength (SSS), using a cutoff of SSS
= 0.80 (or 80\% of the population signal). Thus, for this analysis only,
we defined chronology start dates as the year the SSS exceeded 0.80 or
two years after the start of the climate record, whichever came later.
SSS exceeded 0.80 well before the start of the 1901 start of climate
records for PSME (1800s), ABAL (1700), and PIMA (1850s). For LITU , SSS
reached 0.8 with 11 trees in 1919, which we used as the start date for
this series. We note that these start date criteria differ from those
used in the main analysis (Table S3), which had earlier start dates
because the analysis was not constrained by a need to represent the full
population signal. End dates were defined as the last full year prior to
sampling (Table S3).

@@

Our analysis included two sites where climate change has had pronounced
effects on tree growth: Scotty Creek, NW Territories, Canada (SC) and
Little Tesuque, New Mexico, USA (LT). At SC, {[}temperatures have
increased by X \(^\circ\) over X years{]}\ldots, resulting in negative
growth trends in basal area index (\(BAI\)) starting around 1950 and
significant growth declines since 1970 in 56\% of trees (Sniderhan \&
Baltzer, 2016). At LT, \emph{(drought has increased dramatically)},
resulting in many missing rings in recent years.

\emph{This is in process. We will try and compare 3 methods: (1) our
standard approach, (2) detrending the climate variables (\#53), (3)
applying the climwin step only for older records--before the most rapid
climate change. We will work with SC and LT researchers to determine
which makes most sense, and use that as the main approach for these
sites.}

@@

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{l>{\raggedright\arraybackslash}p{3cm}rr>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{1.5cm}llll}
\toprule
site code & site name & latitude & longitude & elevation (m.a.s.l.) & cores within ForestGEO plot? & canopy positions & tree statuses & date range & dormant season & months in climwin\\
\midrule
BCI & Barro Colorado Island & 9.15430 & -79.8461 & 120-160 & no & canopy & live, dead & 1931-2014 & Nov-Apr & pOct-cDec\\
HKK & Huai Kha Khaeng & 15.63240 & 99.2170 & 549-638 & no & all & live & 1903-2011 & Nov-Apr & pOct-cDec\\
LT & Little Tesuque & 35.73838 & -105.8382 & 2682 & n.a. & all & live & 1903-2018 &  & pMay-cAug\\
CB & Utah Forest Dynamics Plot & 37.66150 & -112.8525 & 3020-3169 & yes &  & live & 1903-2007 &  & pMay-cAug\\
SCBI & Smithsonian Conservation Biology Institute & 38.89350 & -78.1454 & 273-338 & yes & all & live, dead & 1903-2017 & Oct-Apr & pMay-cAug\\
\addlinespace
LDW & Lilly Dickey Woods & 39.23590 & -86.2181 & 230-303 &  & canopy & live, dead & 1903-2019 &  & pMay-cAug\\
HF & Harvard Forest & 42.53880 & -72.1755 & 340-368 & yes & all & live, dead & 1903-2014 &  & pMay-cAug\\
NE & Niobrara/Halsey & 42.78000 & -100.0210 & 644-702 & some & canopy & live &  & Oct-Apr & pMay-cAug\\
ZOF & $\v{Z}$of$\'{i}$n Forest Dynamics Plot & 48.66380 & 14.7073 & 736-829 & some & all & live, dead & 1903-2013 & Oct-Mar & pMay-cAug\\
SC & Scotty Creek & 61.30000 & -121.3000 & 258-274 & no & all & live, dead & 1903-2013 &  & pMay-cAug\\
\bottomrule
\end{tabular}}
\end{table}

@@

\emph{(\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/72}{ISSUE
\#72 in ForestGEO-climate-sensitivity})}

\emph{NOTE: bark.allometry field is not yet right-- we will have just
one latin name per site, corresponding to allometries in Table S4. But
it does give correct info for what is currently applied. We also intend
to find and apply more allometries.}

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lllllll>{\raggedright\arraybackslash}p{5cm}}
\toprule
\textbf{species code} & \textbf{family} & \textbf{latin name} & \textbf{sites sampled} & \textbf{leaf type} & \textbf{leaf phenology} & \textbf{light requirements} & \textbf{bark allometry}\\
\midrule
ABAL & Pinaceae & Abies alba & ZOF & needleleaf & evergreen & shade-tolerant & neglected in Zofin\\
\addlinespace
ABBI & Pinaceae & Abies bifolia & CB & needleleaf & evergreen &  & neglected in CedarBreaks\\
\addlinespace
ACRU & Sapindaceae & Acer rubrum & HF & broadleaf & deciduous (cold) &  & acru in HarvardForest\\
\addlinespace
ACSA & Sapindaceae & Acer saccharum & LDW & broadleaf & deciduous (cold) &  & acru in LillyDickey, acru in LillyDickey\\
\addlinespace
AFXY & Fabaceae & Afzelia xylocarpa & HKK & broadleaf & deciduous (drought) &  & neglected in HKK\\
\addlinespace
BEAL & Betulaceae & Betula alleghaniensis & HF & broadleaf & deciduous (cold) &  & Betula alleghaniensis in HarvardForest\\
\addlinespace
BEPA & Betulaceae & Betula papyrifera & NE & broadleaf & deciduous (cold) &  & Betula papyrifera in Nebraska\\
\addlinespace
CACO & Juglandaceae & Carya cordiformis & SCBI & broadleaf & deciduous (cold) &  & caco in SCBI\\
\addlinespace
CAGL & Juglandaceae & Carya glabra & SCBI & broadleaf & deciduous (cold) &  & cagl in SCBI\\
\addlinespace
CAOV & Juglandaceae & Carya ovata & LDW & broadleaf & deciduous (cold) &  & cagl in LillyDickey\\
\addlinespace
CAOVL & Juglandaceae & Carya ovalis & SCBI & broadleaf & deciduous (cold) &  & caovl in SCBI\\
\addlinespace
CATO & Juglandaceae & Carya tomentosa & SCBI & broadleaf & deciduous (cold) &  & cato in SCBI\\
\addlinespace
CHTA & Meliaceae & Chukrasia tabularis & HKK & broadleaf & brevi-deciduous (drought) &  & neglected in HKK\\
\addlinespace
FAGR & Fagaceae & Fagus grandifolia & HF, SCBI & broadleaf & deciduous (cold) &  & neglected in HarvardForest, neglected in LillyDickey, neglected in SCBI\\
\addlinespace
FASY & Fagaceae & Fagus sylvatica & ZOF & broadleaf & deciduous (cold) & shade-tolerant & neglected in Zofin\\
\addlinespace
FRAM & Oleaceae & Fraxinus americana & LDW, SCBI & broadleaf & deciduous (cold) &  & Fraxinus ssp. in LillyDickey, fram in SCBI\\
\addlinespace
FRNI & Oleaceae & Fraxinus nigra & SCBI & broadleaf & deciduous (cold) &  & fram in SCBI\\
\addlinespace
JACO & Bignoniaceae & Jacaranada copaia & BCI & broadleaf & deciduous (drought) & light-demanding & JCO in BCI\\
\addlinespace
JUNI & Juglandaceae & Juglans nigra & SCBI & broadleaf & deciduous (cold) &  & juni in SCBI\\
\addlinespace
JUVI & Cupressaceae & Juniperus virginiana & NE &  &  &  & neglected in Nebraska\\
\addlinespace
LITU & Magnoliaceae & Liriodendron tulipifera & LDW, SCBI & broadleaf & deciduous (cold) &  & litu in LillyDickey, litu in LillyDickey, litu in SCBI\\
\addlinespace
MEAZ & Meliaceae & Melia azedarach & HKK & broadleaf & deciduous (drought) & light-demanding & neglected in HKK\\
\addlinespace
PIAB & Pinaceae & Picea abies & HF & needleleaf & evergreen & intermediate & neglected in HarvardForest, neglected in Zofin\\
\addlinespace
PIEN & Pinaceae & Picea engelmannii & CB & needleleaf & evergreen &  & Picea engelmannii in CedarBreaks\\
\addlinespace
PIFL & Pinaceae & Pinus flexilis & CB & needleleaf & evergreen &  & Pinus monticola in CedarBreaks\\
\addlinespace
PILO & Pinaceae & Pinus longaeva & CB & needleleaf & evergreen &  & neglected in CedarBreaks\\
\addlinespace
PIMA & Pinaceae & Picea mariana & SC & needleleaf & evergreen &  & PIMA in ScottyCreek\\
\addlinespace
PIPO & Pinaceae & Pinus ponderosa & NE, LT & needleleaf & evergreen &  & Pinus jeffreyi in Little Tesuque, Pinus jeffreyi in Nebraska\\
\addlinespace
PIPU & Pinaceae & Picea pungens & CB & needleleaf & evergreen &  & neglected in CedarBreaks\\
\addlinespace
PIST & Pinaceae & Pinus strobus & HF, SCBI & needleleaf & evergreen &  & Pinus strobus in HarvardForest, pist in SCBI\\
\addlinespace
PIST2 & Pinaceae & Pinus strobiformis & LT & needleleaf & evergreen &  & Pinus monticola in Little Tesuque\\
\addlinespace
POTR & Salicaceae & Populus tremuloides & CB & broadleaf & deciduous (cold) &  & Populus tremuloides in CedarBreaks\\
\addlinespace
PSME & Pinaceae & Pseudotsuga menziesii & CB & needleleaf & evergreen &  & Pseudotsuga menziesii in CedarBreaks\\
\addlinespace
QUAL & Fagaceae & Quercus alba & LDW, SCBI & broadleaf & deciduous (cold) &  & qual in LillyDickey, qual in SCBI\\
\addlinespace
QUMO & Fagaceae & Quercus montana & LDW, SCBI & broadleaf & deciduous (cold) &  & qupr in LillyDickey, qupr in SCBI\\
\addlinespace
QURU & Fagaceae & Quercus rubra & HF, LDW, SCBI & broadleaf & deciduous (cold) &  & quru in HarvardForest, Quercus rubra in LillyDickey, quru in SCBI\\
\addlinespace
QUVE & Fagaceae & Quercus velutina & LDW, SCBI & broadleaf & deciduous (cold) &  & quve in LillyDickey, quve in SCBI\\
\addlinespace
TEPA & Burseraceae & Tetragastris panamensis & BCI & broadleaf & evergreen & shade-tolerant & TPA in BCI\\
\addlinespace
TOCI & Meliaceae & Toona ciliata & HKK & broadleaf & deciduous (drought) &  & neglected in HKK\\
\addlinespace
TRTU & Meliaceae & Trichilia tuberculata & BCI & broadleaf & evergreen & shade-tolerant & TTU in BCI\\
\addlinespace
TSCA & Pinaceae & Tsuga canadensis & HF & needleleaf & evergreen &  & Tsuga canadensis in HarvardForest\\
\bottomrule
\end{tabular}}
\end{table}

@@

\emph{(\href{https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/73}{ISSUE
\#73 in ForestGEO-climate-sensitivity})}

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{llrrrrlll}
\toprule
\textbf{site} & \textbf{species code} & \textbf{n trees all} & \textbf{n cores all} & \textbf{n trees dbh} & \textbf{n cores dbh} & \textbf{dbh range sampled} & \textbf{dbh range reconstructed*} & \textbf{date range}\\
\midrule
BCI & JACO & 12 & 18 & 11 & 17 & 30.2-63.5 & 2.6-56.4 & 1931-2014\\
\addlinespace
BCI & TEPA & 18 & 29 & 17 & 26 & 22.1-59.5 & 2.7-49.4 & 1931-2014\\
\addlinespace
BCI & TRTU & 23 & 37 & 20 & 31 & 20.7-43.6 & 4.8-41.5 & 1931-2014\\
\addlinespace
CB & ABBI & 22 & 41 & 20 & 37 & 13.9-54.2 & 0.9-46.4 & 1903-2000\\
\addlinespace
CB & PIEN & 12 & 21 & 10 & 15 & 14-54.9 & 0.9-33.1 & 1903-2000\\
\addlinespace
CB & PIFL & 13 & 22 & 12 & 21 & 17.6-64.1 & 1.5-47.5 & 1903-1998\\
\addlinespace
CB & PILO & 11 & 16 & 8 & 12 & 45.8-121.6 & 35.4-118.5 & 1903-1999\\
\addlinespace
CB & PIPU & 15 & 28 & 15 & 28 & 22.4-50.8 & 9.5-48.4 & 1903-2000\\
\addlinespace
CB & POTR & 17 & 27 & 17 & 26 & 23.6-47.6 & 4.3-35.4 & 1903-2000\\
\addlinespace
CB & PSME & 11 & 20 & 10 & 18 & 20.7-64.2 & 0.5-41.5 & 1903-1999\\
\addlinespace
HF & ACRU & 18 & 59 & 18 & 59 & 10.1-22.1 & 0.9-20.4 & 1903-2013\\
\addlinespace
HF & BEAL & 13 & 44 & 13 & 44 & 10.2-37.9 & 0-17.2 & 1904-2013\\
\addlinespace
HF & QURU & 74 & 180 & 73 & 177 & 19.5-53 & 1.1-48.3 & 1903-2014\\
\addlinespace
HF & TSCA & 32 & 83 & 32 & 71 & 10.6-37 & 0-28.4 & 1923-2014\\
\addlinespace
HKK & AFXY & 39 & 127 & 39 & 127 & 20.1-98.7 & 0.1-81.4 & 1903-2011\\
\addlinespace
HKK & CHTA & 28 & 70 & 28 & 70 & 16-64.6 & 0.2-59.5 & 1904-2010\\
\addlinespace
HKK & MEAZ & 46 & 130 & 46 & 130 & 25.6-98.1 & 3.8-80.3 & 1914-2011\\
\addlinespace
HKK & TOCI & 45 & 143 & 45 & 143 & 16.6-116.4 & 1.7-80.5 & 1903-2011\\
\addlinespace
LDW & ACSA & 35 & 66 & 34 & 64 & 9-64.6 & 0-52.4 & 1903-2019\\
\addlinespace
LDW & CAOV & 9 & 18 & 8 & 15 & NA-NA & 1.4-37.4 & 1903-2013\\
\addlinespace
LDW & LITU & 15 & 28 & 14 & 25 & NA-NA & 1.2-69.4 & 1903-2019\\
\addlinespace
LDW & QUAL & 10 & 20 & NA & NA & NA & NA & 1903-2013\\
\addlinespace
LDW & QUMO & 10 & 20 & 8 & 16 & NA-NA & 1.1-52.4 & 1903-2013\\
\addlinespace
LDW & QUVE & 9 & 18 & NA & NA & NA & NA & 1903-2013\\
\addlinespace
NE & BEPA & 28 & 84 & 28 & 84 & NA-NA & 0-25.5 & 1903-1995\\
\addlinespace
NE & JUVI & 29 & 60 & 29 & 60 & 16.6-21.5 & 0.3-18.4 & 1938-1994\\
\addlinespace
NE & PIPO & 71 & 134 & 70 & 129 & 19.3-64 & 0-41.1 & 1934-1994\\
\addlinespace
LT & PIPO & 10 & 20 & 10 & 20 & 23.2-52.8 & 14.4-48.4 & 1903-2018\\
\addlinespace
LT & PIST2 & 7 & 14 & 7 & 14 & 25.7-39.8 & 3.9-34.5 & 1903-2018\\
\addlinespace
SCBI & CACO & 15 & 15 & 14 & 14 & 10.62-38.52 & 1.7-32.2 & 1903-2015\\
\addlinespace
SCBI & CAGL & 39 & 39 & 36 & 36 & 10.28-52.31 & 1.6-49.3 & 1903-2015\\
\addlinespace
SCBI & CAOVL & 25 & 25 & 24 & 24 & 15.11-60.32 & 2.6-47.2 & 1903-2015\\
\addlinespace
SCBI & CATO & 15 & 15 & 14 & 14 & 12.86-35.95 & 3.7-28.4 & 1903-2015\\
\addlinespace
SCBI & FAGR & 76 & 76 & 74 & 74 & 10.05-41.02 & 0.1-41.2 & 1920-2009\\
\addlinespace
SCBI & FRAM & 66 & 66 & 61 & 61 & 8.11-94.73 & 0.1-84.4 & 1903-2016\\
\addlinespace
SCBI & FRNI & 12 & 12 & 12 & 12 & 11.04-39.2 & 0.5-27.3 & 1903-1996\\
\addlinespace
SCBI & JUNI & 30 & 30 & 28 & 28 & 20.4-76.19 & 5.6-59.5 & 1903-2010\\
\addlinespace
SCBI & LITU & 106 & 106 & 104 & 104 & 10-91.42 & 0.1-81.1 & 1903-2010\\
\addlinespace
SCBI & PIST & 36 & 36 & 36 & 36 & 13.92-50.96 & 1.6-45.2 & 1931-2010\\
\addlinespace
SCBI & QUAL & 66 & 66 & 66 & 66 & 11.4-76.73 & 0.3-70.4 & 1903-2009\\
\addlinespace
SCBI & QUMO & 67 & 67 & 67 & 67 & 10.22-84.59 & 0.3-69.5 & 1903-2017\\
\addlinespace
SCBI & QURU & 70 & 70 & 70 & 70 & 11.07-87.65 & 2.5-79.2 & 1903-2016\\
\addlinespace
SCBI & QUVE & 81 & 81 & 81 & 81 & 16.02-82.33 & 0.5-78.4 & 1903-2009\\
\addlinespace
SC & PIMA & 442 & 442 & 101 & 101 & 7-14.9 & 0.5-12.5 & 1903-2013\\
\addlinespace
ZOF & ABAL & 55 & 55 & 50 & 50 & 41-121 & 21.3-108.5 & 1903-2011\\
\addlinespace
ZOF & FASY & 1369 & 1369 & 1358 & 1358 & NA-NA & 0.1-115.3 & 1903-2013\\
\addlinespace
ZOF & PCAB & 644 & 644 & 599 & 599 & NA-NA & 0.5-126.4 & 1903-2011\\
\bottomrule
\end{tabular}}
\end{table}

*Maximum reconstructed \(DBH\)'s analyzed are less than maximum sampled
\(DBH\)'s because we discard size ranges with \textless{} 3 conspecific
trees.

@@

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{}llrl>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{6cm}}
\toprule
\textbf{species} & \textbf{equation} & \textbf{n} & \textbf{DBH.range.cm} & \textbf{site} & \textbf{source}\\
\midrule
\em{Acer rubrum} & $bark.mm = 0.619 * log(dbh.cm + 1)$ & 10 & 8.2-39.6 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Betula alleghaniensis} & $bark.mm = (0.15 + 0.03 * dbh.cm)/2$ & NA &  & North America average of its range & Miles, Patrick D.; Smith, W. Brad. 2009.\\
\addlinespace
\em{Betula papyrifera} & $bark.mm = (0.13 + 0.05 * dbh.cm)/2$ & NA &  & North America average of its range & Miles, Patrick D.; Smith, W. Brad. 2009.\\
\addlinespace
\em{Carya cordiformis} & $bark.mm = 0.793 * log(dbh.cm + 1)$ & 9 & 5.9-68.2 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Carya ovalis} & $bark.mm = 1.531 * log(dbh.cm + 1)$ & 8 & 6.4-63.1 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Carya ovata} & $bark.mm = 1.035 * log(dbh.cm + 1)$ & 8 & 19.1-78 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Carya tomentosa} & $bark.mm = 1.105 * log(dbh.cm + 1)$ & 8 & 5-57.3 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Fraxinus americana} & $bark.mm = 2.223 * log(dbh.cm + 1)$ & 9 & 6.1-94.2 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Fraxinus americana} & $bark.mm = (0.38 + 0.05 * dbh.cm)/2$ & NA &  & North America average of its range & Miles, Patrick D.; Smith, W. Brad. 2009.\\
\addlinespace
\em{Jacaranada copaia} & $bark.mm = 2.993 * log(dbh.cm + 1)$ & 5 & 45.6-75 & Panama & Raquel Alfaro-S$\acute{a}$nchez (unpublished data)\\
\addlinespace
\em{Juglans nigra} & $bark.mm = 2.107 * log(dbh.cm + 1)$ & 9 & 13.6-85.4 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Liriodendron tulipifera} & $bark.mm = 1.637 * log(dbh.cm + 1)$ & 9 & 27.5-136.5 & SCBI & Anderson-Teixeira et al. (2015)\\
\addlinespace
\em{Picea engelmannii} & $bark.mm = (0.15 + 0.04 * dbh.cm)/2$ & NA &  & North America average of its range & Miles, Patrick D.; Smith, W. Brad. 2009.\\
\addlinespace
